The \tsc{r}-pronoun and postposition \tit{waar-mee} in Dutch

### 2020-04-16 presentation nanolab/presentation.tex
presentation thursday
am ende noch ein bisschen

\exg. *Ik schilder \tbf{met} \tbf{t}.\\
 I paint with it\\
 `I am painting with it.'\label{ex:neemett}

@@

\exg. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint there -with\\
 `I am painting with it.'\label{ex:jarmee}

@@

*\tit{met t} vs. \tit{r mee}

@@

\begin{itemize}
	\item \tit{met} becomes \tit{mee} \pause
	\item position of adposition differs \pause
	\item \tsc{r}-pronouns and locatives are syncretic
\end{itemize}

@@

@@

\exg. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint there -with\\
 `I am painting with it.'\label{ex:preper}

\exg. *Ik schilder \tbf{mee} \tbf{r}.\\
 I paint with there\\
 `I am painting with it.'\label{ex:erprep}

@@

\ex.
\ag. Ik zit \tbf{r} \tbf{op}.\\
 I sit there on\\
 `I am sitting on it.
\bg. *Ik zit \tbf{op} \tbf{t}.\\
 I sit on it\\
 `I am sitting on it.

\ex.
 \ag. Hij zwemt \tbf{r} \tbf{in}.\\
  he swims it-in\\
  `He is swimming there it.'
 \bg. *Hij zwemt \tbf{in} \tbf{t}.\\
  he swims in it\\
  `He is swimming in it.'

@@

\exg. \tbf{Waar} \tbf{-mee} schilder jij?\\
 where with paint you with\\
 `What are you painting with?'\label{ex:waar-mee}

\ex.\label{ex:headed}
\ag. Ik schilder met de kwast \tbf{waar} \tbf{-mee} jij ook schildert.\\
 I paint with the brush where with you also paint\\
 `I am painting with the brush that you are painting with too.'
\bg. *Ik schilder met de kwast \tbf{met} \tbf{wat} jij ook schildert.\\
 I paint with the brush with what you also paint\\
 `I am painting with the brush that you are painting with too.'

\ex.\label{ex:headless}
\ag. Ik schilder \tbf{waar} \tbf{-mee} jij ook schildert.\\
 I paint where with you also paint\\
 `I am painting with what you are painting with too.'
\bg. *Ik schilder \tbf{met} \tbf{wat} jij ook schildert.\\
 I paint with what you also paint\\
 `I am painting with what you are painting with too.'

\ex.
\ag. Ik heb gekocht \tbf{waar} \tbf{-mee} jij schildert.\\
 I have bought where with you paint\\
 `I bought what you are painting with.'\label{ex:gekochtwaar-mee}
\bg. *Ik heb gekocht \tbf{met} \tbf{wat} jij schildert.\\
 I have bought with what you paint\\
 `I bought what you are painting with.'\label{ex:gekochtmetwat}

\ex.\label{ex:schildermet}
\ag. *Ik schilder \tbf{waar} \tbf{-mee} jij hebt gekocht.\\
 I paint where with you have bought\\
 `I paint with what you bought.'\label{ex:schilderwaar-mee}
\bg. Ik schilder \tbf{met} \tbf{wat} jij hebt gekocht.\\
 I paint with what you have bought\\
 `I paint with what you bought.'\label{ex:schildermetwat}

\ex.
\ag. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint there with\\
 `I am painting with it.'\label{ex:const1}
\bg. \tbf{Waar} \tbf{-mee} schilder jij?\\
where with paint you with\\
 `What are you painting with?'\label{ex:const2}
\bg. Ik schilder met de kwast [\tbf{waar} \tbf{-mee} jij ook schildert].\\
 I paint with the brush where with you also paint\\
 `I am painting with the brush that you are painting with too.'\label{ex:const3}
\bg. Ik schilder [\tbf{waar} \tbf{-mee} jij ook schildert].\\
 I paint where with you also paint\\
 `I am painting with what you are painting with too.'\label{ex:const4}

@@

\exg. Ik heb gekocht [\tbf{waar} \tbf{-mee} jij schildert].\\
 I have bought where with you paint\\
 `I bought what you are painting with.'\label{ex:mismatchwaar-mee}

@@

\exg. [\tbf{Met} [\tbf{wat} [voor [potloden]]] teken jij?\\
 with what for pencils draw you\\
 `What kind of pencils do you with?'\label{ex:watwasfur}
\exg. Ik wil graag thee [\tbf{met} [\tbf{wat} [suiker]]].\\
 I want please tea with some sugar\\
 `I would like to have tea with some sugar.'\label{ex:watindef}

@@

\ex.\label{ex:decompose}
\ag. w -aa -r -mee\\
\tsc{w} -\tsc{deix} -\tsc{loc} -\tsc{ins}\\
\bg. met w -a -t\\
\tsc{ins} \tsc{w} -\tsc{deix} -\tsc{n.sg}\\
\z.
\z.

@@

\begin{frame}{ \tit{w} and \tit{a}}

@@

\begin{frame}{\tit{t}}

@@

\begin{frame}{\tit{r}}

\exg. Ik ben r al geweest.\\
 I am there already been\\
 `I have already been there.'

@@

\begin{frame}{\tit{mee}}

@@

\begin{frame}{\tit{met}}

@@

@@

@@

@@

\ex.
\a. \tit{met} --- \tit{mee}
\b. \tit{tot} --- \tit{toe}

\ex. CVt
\pause

\ex.
\a. \tit{in} --- \tit{in}
\b. \tit{van} --- \tit{van}
\b. \tit{op} --- \tit{op}

\appendix

\begin{frame}{References}

	\scriptsize{
		\printbibliography
	}

>>>>

@@

\exg. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint \tsc{r} -with\\
 `I am painting with it.'\label{ex:jarmee}

@@

\centering{
*\tit{met t} \hspace{1em} | \hspace{1em} \tit{r mee}}

@@

\begin{itemize}
	\item positions are reversed \pause
	\item \tit{met} becomes \tit{mee} \pause
	\item \tit{t} becomes \tit{r}, which is syncretic with locative
\end{itemize}

\tit{rmee} (and not \tit{met t}) appears when all features form a proper constituent

@@

\ex. rmee ⇒ waarmee

\ex. met t ⇒ met wat

\pause
\vspace{1em}

\exg. \tbf{Waar} \tbf{-mee} schilder jij?\\
 \tsc{r} with paint you with\\
 `What are you painting with?'\label{ex:waar-mee}

@@

\ex.\label{ex:headed}
\ag. Ik schilder met de kwast \tbf{waar} \tbf{-mee} jij ook schildert.\\
 I paint with the brush \tsc{r} with you also paint\\
 `I am painting with the brush that you are painting with too.'
\bg. *Ik schilder met de kwast \tbf{met} \tbf{wat} jij ook schildert.\\
 I paint with the brush with what you also paint\\
 `I am painting with the brush that you are painting with too.'

\ex.\label{ex:headless}
\ag. Ik schilder \tbf{waar} \tbf{-mee} jij ook schildert.\\
 I paint \tsc{r} with you also paint\\
 `I am painting with what you are painting with too.'
\bg. *Ik schilder \tbf{met} \tbf{wat} jij ook schildert.\\
 I paint with what you also paint\\
 `I am painting with what you are painting with too.'

@@

\ex.
\ag. Ik heb gekocht \tbf{waar} \tbf{-mee} jij schildert.\\
 I have bought \tsc{r} with you paint\\
 `I bought what you are painting with.'\label{ex:gekochtwaar-mee}
\bg. *Ik heb gekocht \tbf{met} \tbf{wat} jij schildert.\\
 I have bought with what you paint\\
 `I bought what you are painting with.'\label{ex:gekochtmetwat}

\ex.\label{ex:schildermet}
\ag. *Ik schilder \tbf{waar} \tbf{-mee} jij hebt gekocht.\\
 I paint \tsc{r} with you have bought\\
 `I paint with what you bought.'\label{ex:schilderwaar-mee}
\bg. Ik schilder \tbf{met} \tbf{wat} jij hebt gekocht.\\
 I paint with what you have bought\\
 `I paint with what you bought.'\label{ex:schildermetwat}

@@

\ex.
\ag. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint \tsc{r} with\\
 `I am painting with it.'\label{ex:const1}
\bg. \tbf{Waar} \tbf{-mee} schilder jij?\\
 \tsc{r} with paint you with\\
 `What are you painting with?'\label{ex:const2}
\bg. Ik schilder met de kwast [\tbf{waar} \tbf{-mee} jij ook schildert].\\
 I paint with the brush \tsc{r} with you also paint\\
 `I am painting with the brush that you are painting with too.'\label{ex:const3}
\bg. Ik schilder [\tbf{waar} \tbf{-mee} jij ook schildert].\\
 I paint \tsc{r} with you also paint\\
 `I am painting with what you are painting with too.'\label{ex:const4}

@@

\exg. Ik heb gekocht [\tbf{waar} \tbf{-mee} jij schildert].\\
 I have bought \tsc{r} with you paint\\
 `I bought what you are painting with.'\label{ex:mismatchwaar-mee}

@@

\exg. [\tbf{Met} [\tbf{wat} [voor [potloden]]] teken jij?\\
 with what for pencils draw you\\
 `What kind of pencils are you drawing with?'\label{ex:watwasfur}

\exg. Ik wil graag thee [\tbf{met} [\tbf{wat} [suiker]]].\\
 I want please tea with some sugar\\
 `I would like to have tea with some sugar.'\label{ex:watindef}

@@

\ex. w -aa -r -mee

\ex. met w -a -t

@@

@@

@@

\exg. Ik ben r al geweest.\\
 I am \tsc{r} already been\\
 `I have already been there.'

@@

@@

@@

	show fseq

@@

	\begin{itemize}
		\item rmee is nothing special anymore, just a result of spellout
		\item r is ambiguous between thing and location, r is a caseless base
		\item met and mee are separate lexical entries, so the system allows for them to phonologically differ
		\item mee is a postposition, met is a preposition
	\end{itemize}

	what happens to \tit{met wat}?

	--give a tree of halfway there--

\exg. Ik schilder \tbf{r} \tbf{-mee}.\\
 I paint \tsc{r} -with\\
 `I am painting with it.'\label{ex:preper}

gender and number between sigma and case -- make tree

list of postpositions and prepositions
maybe i have a tree

\ex.
\ag. Ik zit \tbf{r} \tbf{op}.\\
 I sit \tsc{r} on\\
 `I am sitting on it.
\bg. *Ik zit \tbf{op} \tbf{t}.\\
 I sit on it\\
 `I am sitting on it.

\ex.
 \ag. Hij zwemt \tbf{r} \tbf{in}.\\
  he swims it-in\\
  `He is swimming in it.'
 \bg. *Hij zwemt \tbf{in} \tbf{t}.\\
  he swims in it\\
  `He is swimming in it.'

verbal particles

\ex. CVt
\a. \tit{met} --- \tit{mee}
\b. \tit{tot} --- \tit{toe}

	repeat example i always use

@@

	look at story again

@@
